---
title: "Assignment 4"
author: "Daniel Stewart"
date: "October 7, 2019"
output: html_document
---

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

Installed Packages:

  + install.packages("tidyverse")
  + install.packages("gapminder")
  + install.packages("ggplot2")
  + install.packages("tidyr")
  + install.packages("readr")

```{r setup, include=FALSE, message=F}
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(readr))
```

## Exercise 1: Univariate Data Reshaping
### Univariate Option 1

#### Part 1: Life Expectancy of Scandinavian Countries (1952-2007)
```{r}
Scandinavian_LifeExp <- gapminder %>%
  filter(country == "Denmark" | country == "Sweden" | country == "Finland" | country == "Iceland") %>%
  pivot_wider(id_cols = year,
  names_from = country,
  values_from = lifeExp)
  Scandinavian_LifeExp
  
#Note: in the filtering process I had to identify the "country" field for each country.
```

#### Part 2: Scatterplot of Scandinavian Country Life Expectancies (1952-2007)
```{r}
PlotData <- reshape2::melt(Scandinavian_LifeExp, id.vars = "year") %>%
  rename(Country = variable)

ggplot(PlotData, aes(year, value, col=Country)) +
  geom_point () + 
  facet_wrap(~Country) +
  xlab("Year") +
  ylab("Life Expectancy") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

#I tried to plot the outputs of Part 1 directly, but it was extremely difficult due to the multiple columns. I initially solved this by using "geom_point" separately for each country, but this made it impossible to generate an appropriate legend. In the end I opted to melt the columns into one, allowing me to make these plots in a satisfactory manner. 
```
#### Part 3: Relengthened Data

```{r}
Relengthened_data <- Scandinavian_LifeExp %>%
  pivot_longer(cols = c(Denmark, Finland, Iceland, Sweden),
               names_to = "country",
               values_to = "life_expectancy")
Relengthened_data
```

## Exercise 2: Multivariate Data Reshaping
### Multivariate Option 1

#### Part 1: Life Expectancy and GDP per Capita of Three Randomly-Selected Countries
```{r}
  widened_multivariate <- gapminder %>%
  filter(country == "Eritrea" | country == "Jamaica" | country == "Vietnam") %>%
  pivot_wider(id_cols = year,
  names_from = country,
  values_from = c(lifeExp, gdpPercap))
  widened_multivariate
```
#### Part 2: Relengthened Data
```{r}
Relengthened_multivariate <- widened_multivariate %>%
  pivot_longer(cols = c(-year),
               names_to = c("metric", "country"),
               names_sep = ("_"),
              values_to = "measurement") 

Relengthened_multivariate
#At this stage lifeexp and gdpPercap are combined under the "metric" column. This is as "narrow" as I could make it. 

Relengthened_multivariate %>%
  pivot_wider(names_from = metric,
              values_from = measurement)

#This code returns it to its original format



```

## Exercise 3: Table Joins
```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv") %>%
  rename(name = guest)
#I had to rename the "guest" column to make it compatible with guest tribble
```


#### Part 3.1: Adding an e-mail column to the guest tibble 
```{r}
emailjoin <- guest %>% 
  full_join(email, by = c("name"))
emailjoin
```
#### Part 3.2: People not on the guest list who we have emails for 
```{r}
emailjoin %>%
  filter(is.na(party)) %>%
  filter(!is.na(email))

#I interpreted NA values across all attendance fields as == not being on the guest list.


```
#### Part 3.3: Guestlist that includes everyone we have emails for 

```{r}
masteremail <- emailjoin %>% 
  semi_join(email, by ="name")
  masteremail
```


